FAIR Theory: Applying Open Science Principles to the Construction and Iterative Improvement of Scientific Theories

### 2024-12-12 manuscript.Rmd
Update manuscript.Rmd

The present paper addresses the lack of open science methods for theory development and improvement by applying the FAIR principles to scientific theories.
Merely publishing theory does not make it open;
to be open, theory should adhere to established open science standards.
The FAIR principles, initially introduced for open research data,
have since been applied to other information artifacts [@lamprechtFAIRPrinciplesResearch2019].
We apply the FAIR principles to digital representations of theory,
introducing a FAIR metadata format to make theories *Findable* via a DOI,
*Accessible* in a machine- and human-readable filetype,
*Interoperable* within the data analysis environment,
and *Reusable* in the practical and legal sense, so that they may be improved over time.
Following Lamprecht and colleagues,
we adapt the FAIR principles for theory, see Table \@ref(tab:tabfair).
We reflect on the necessary changes (which are minor),
as well as on the current state and future of FAIR theory in the social sciences.
The resulting principles provide guidance for instantiating theory as a FAIR information artifact,
and we provide worked examples to encourage their adoption.

>>>>

The present paper addresses the lack of open science methods for theory development and suggests an improvement of the state of affairs by applying the FAIR principles to scientific theories.
Merely publishing theory -- as in a classic research article -- does not make it open;
to be open, theory should adhere to established open science standards.
<!--AB  I deleted this sentence because it was already in the introduction -->
We apply the FAIR principles to digital representations of theory,
introducing a FAIR metadata format to make theories *Findable* via a DOI,
*Accessible* in a machine- and human-readable filetype,
*Interoperable* within the data analysis environment,
and *Reusable* in the practical and legal sense, so that they may be improved over time -- at best, in a participative process.
Digital representations of theory intentionally is a broad term, particularly including textual representations of a given theory, as well as formal representations, such as mathematical notation, algorithmic pseudo code, or a set of logical clauses. <!-- AB I think some initial word on what is covered is needed here; otherwise readers may wonder what we mean to cover--> 
Following the original proposal of Lamprecht and colleagues,
we adapt the FAIR principles for theory, see Table \@ref(tab:tabfair).
We reflect on the necessary changes (which are minor),
as well as on the current state and future of FAIR theory in the social sciences.
The resulting principles provide guidance for instantiating theory as a FAIR information artifact,
and we provide worked examples to encourage their adoption.
